---
title: "Reproducible Papers"
author: "Sam Dunn"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


We will use {rrtools} to generate reproducible manuscripts and even ombine al elements of a paper into a package.  use the [@citation] format to access the .bib file in the wd.  Can also control reference styles programmatically.


## Mathematical Notation
\[\sum{i=}^{N}{log(i)+ \frac{\omega}[x]}\]  Write in \[latex\] or mathjacks


use kable for satic figure...or htmltable for pretty formatting

Create microchunks using single back-ticks `1+1`
This allows us to embed results in text directly!

## rrTools
Uility package that helps set up R packages for publication. We will primarily use the command [createcompendium].  rrtools isn' on CRAN yet, so we can use the devtols package.
```{r}
# install.packages("devtools")
#devtools::install_github("benmarwick/rrtools")
library(rrtools)

```

Be carefult o set th wd to a location where you want this new package to be placed.  I left it in place
```{r}
#create_compendium("testpaper2")  one time operation.  generates whole directory to support paper writing!
```


